\acknowledgments{
I would like to thank my supervisors Prof. Ramchandra Phawade IIT Dharwad and Prof. S. Sheerazuddin, NIT Calicut for their guidance, support and patience while I completed my dissertation. I thank Prof. Phawade for guiding me in the world of concurrent systems. I thank Prof. S. Sheerazuddin for the Formal methods course he taught me, which planted the seeds of this work and the weekend discussions at CMI, even when I was working in industry. 

I am grateful for the support of my RPC committee Prof. Sreejith, IIT Goa,  Prof. A. Bhaskar, BITS Goa, whose criticisms have vastly shaped this work.

Special thanks to my collaborators, Soumodev Mal and Dr. Francesco Di Cosmo, from whom I learnt a lot academically and also that research can indeed be enjoyable. I am fortunate to have Deepan Raj, for the programming sessions, for being my technical helpline and emotional support especially during COVID times and for teaching me resilience. I look forward to continue working with each of you. 

I thank Prof. Ramanujam for the timely technical discussions which helped me better understand the subject, for nudging me to finally write the dissertation and for graciously hosting me at Azim Premji University multiple times.

I thank Prof. Aiswarya for the insightful discussions and for hosting me at CMI in the summers.

Thanks to Pranav Ashok and the team at Delab Research, it was fun to collaborate on the SAT/SMT/LP solving with you.

Thanks are due to the administration and staff of IIT Dharwad and CSE department faculty, who contributed logistically. Heartfelt thanks to Prof. Koteswara Rao Kondepu, who facilitated my research visit to Rome and to Prof. Rajshekar K who motivated me, especially during the low phases.

I thank Prof. Kumar Madhukar for discussions on building the verification tools, Prof. Madhavan Mukund for shaping my research statement in the initial phase of my PhD journey, Prof. Anantha Padmanabha for discussions on logic and the wonderful company, Sahil Mhaskar for being a sounding board for math, automata, logic ideas and for the friendship, Aditya Shrotri for the guidance and support; Prof. Sandra Kiefer, Prof. Prakash Panangaden and the Formal Methods India community, Women in Logic community at large, for the early feedback on my work and encouragement throughout my research journey.

I thank the PhD students at IIT Dharwad, NIT Calicut and CMI for the camaraderie; and the colleagues and friends who checked in on me periodically, hosted me and ensured that I did not give up.

I am grateful to my family for their love, strength and prayers, and in particular, my mother for instilling in me, the joy of learning.

\vspace{2ex}

\begin{flushright}
\makebox[4cm]{\hrulefill} \\
Tephilla Prince
\end{flushright}
~                                                                        
~                                                                        
~                                                                        
~                                                                        
~                                   

}
