\documentclass{article}
\usepackage{graphicx} % Required for inserting images
\usepackage{todonotes}
\title{Brief Bio-Data of potential Examiners to evaluate Tephilla Prince's PhD dissertation}
\author{ }
\date{ }

\begin{document}

\maketitle

\section{List of Non-Indian Reviewers:}
\begin{enumerate}
    \item Lo\"{i}c Helouet 
    \item[] Designation: Directeur de recherche (Research Director), INRIA  Rennes
    \item[] Webpage: https://people.irisa.fr/Loic.Helouet/
    \item[] Email:  loic.helouet@inria.fr 
    \item[] Contact Number: +33 2 99 84 75 90
    \item[] Education: PhD, IRISA/INRIA, France 
    \item[] Research Interests: Distributed systems : Modeling, semantics, verification
True concurrency models (Petri nets, partial orders, scenarios), Robustness of concurrent and timed specifications. Formal methods for Data centric workflows.
    

    \item Daniel Moldt
    \item[] Designation: Wissenschaftlicher Mitarbeiter, Research Associate, Head of the Laboratory for Agent and Organization-Oriented Systems (LAOS), Univerisity of Hamburg, Germany 
    \item[] Webpage: https://www.inf.uni-hamburg.de/inst/ab/art/people/moldt.html
    \item[] Email: moldt@informatik.uni-hamburg.de
    \item[] Contact Number: +49 40 428 83-2247
    \item[] Education: PhD 
    \item[] Research Interests:
    Petri nets, Business process modeling, Data mining and multi-agent systems, object-oriented analysis of Petri Nets

    \item Michael Blondin
    \item[] Designation: Tenured Associate professor of Computer Science, The Université de Sherbrooke, Québec, Canada 
    \item[] Webpage: https://michaelblondin.com/en/
    \item[] Email: michael.blondin@usherbrooke.ca
    \item[] Contact Number: (+1) 819-821-8000 ext. 66491
    \item[] Education:
    Joint Ph.D from ENS Paris-Saclay and Universite de Montreal
    Postdoctoral fellow at the Technical University of Munich, Germany
    \item[] Research Interests:
    Algorithmic verification, Automata theory, Concurrent and distributed systems, Computational complexity theory and Logic.

    
    

   
    
\end{enumerate}

\section{List of Indian Reviewers:}
\begin{enumerate}
    \item S Akshay
    \item[] Designation: Professor, Indian Institute of Technology Bombay, India 
    \item[] Webpage: https://www.cse.iitb.ac.in/~akshayss/
    \item[] Email: akshayss@ cse.iitb.ac.in
    \item[] Contact Number: +91 (0) 22 2576 7711
    \item[] Education:  Joint PhD degree from Ecole Normale Superieure de Cachan (ENS-Cachan) and Chennai Mathematical Institute (CMI), India. 
    \item[] Research Interests: 
     verification of timed, recursive and distributed systems;
     automated functional synthesis;
     formal modeling and analysis of probabilistic and dynamical systems;
     trust issues and certification in AI models. 

    
    \item Praveen M
    \item[] Designation: Associate Professor at Chennai Mathematical Institute (CMI), Chennai 
    \item[] Webpage: https://www.cmi.ac.in/~praveenm/
    \item[] Email: praveenm@cmi.ac.in
    \item[] Contact Number: +91 44 6748 0966
    \item[] Education: PhD in Theoretical Computer Science at
Institute of Mathematical Sciences. 
    \item[] Research Interests: Computational complexity of modelling and 
            verifying concurrent infinite state systems, logic and parameterized complexity.

    

    \item Rajarshi Ray
    \item[] Designation: Associate Professor at Indian Association for cultivation of Science (IACS) Kolkata. 
    \item[] Webpage: https://iacs.res.in/athusers/index.php?navid=0\&userid=IACS0052
    \item[] Email: rajarshi.ray@iacs.res.in
    \item[] Contact Number: +91 33 2473 4971 (Ext: 2335)
    \item[] Education: PhD in Computer Science from Laboratory Verimag, University of Grenoble. 
    \item[] Research Interests: Formal Verification, Model Checking 
    
\end{enumerate}

\end{document}
